\section*{Acknowledgments}

First and foremost, I am profoundly grateful to God for His blessings, guidance, and strength throughout this journey.

I wish to express my deepest appreciation to my co-supervisors, Prof. Karima Echihabi and Prof. Themis Palpanas, for their unwavering support, encouragement, and insightful guidance. Their expertise and mentorship have been instrumental in the completion of this thesis.

My heartfelt gratitude goes to my family for their unconditional love and support. To my father Ahmed, my beloved wife Diana, my sister Asma, my brothers Anas and Fadlallah—your encouragement and understanding have been my pillars of strength. This work would not have been possible without you.

I dedicate this thesis to the memory of my mother, Habiba, who left us in June 2010. Her love and teachings continue to inspire me every day. I also remember all my departed family members, whose memories motivate me to strive for excellence.

I am sincerely thankful to my dear friends Waiel, Mouna, Mouad, Mohamed, Abdel, Amine, Aissame and all others who have stood by me with their support and camaraderie.

Special thanks to my colleagues from the DMG team at the College of Computing, Mohammed VI Polytechnic University and the DiNo team at the University of Paris Cité. I am grateful for the collaborative environment and the enriching discussions we shared. I also extend my appreciation to the faculty and administrative staff of both universities for their assistance and support.

I would like to acknowledge the UM6P African Supercomputing Center and its support team for providing the computational resources and technical assistance necessary for conducting our experiments and workloads.

Finally, to everyone who has contributed to this journey, whether mentioned here or not, your support has been invaluable, and I am deeply appreciative.
